---
title: "Barplot_Past_Host"
author: "Veronica Radice"
date: "19/04/2021"
output: html_document
---


# Barplots of Differentially expressed genes

## Coral host
### *Porites astreoides*


### Import libraries
```{r}
library(ggplot2)
library(scales)
library(dplyr)
library(plyr)
```

### Resources
https://stackoverflow.com/questions/38268741/geom-bar-ggplot2-stacked-grouped-bar-plot-with-positive-and-negative-values-p


### Set working directory
```{r}
setwd("~/Documents/Rprojects/postdoc\ Rprojects/ODU_postdoc_Rprojects/Paper_Ojo_gene-expression/Porites_astreoides/host/")
```



# Effect of Destination Control

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Past_Host_DE_genes_OC_LC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OC-LC'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))
summary_DEG_1a <- summary_DEG

summary_DEG_1b <- data.frame("group" = "OC-LC", "LFC" = "Down", "n" = 0)
summary_DEG_1 <- rbind(summary_DEG_1a, summary_DEG_1b)

summary_DEG_1[sapply(summary_DEG_1, is.character)] <- lapply(summary_DEG_1[sapply(summary_DEG_1, is.character)], as.factor)

summary_DEG_1
```


###############################################################################

# Effect of Destination Ojo

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Past_Host_DE_genes_OO_LO_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-LO'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_2 <- summary_DEG
summary_DEG_2
```


###############################################################################

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Past_Host_DE_genes_LO_LC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'LO-LC'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_3 <- summary_DEG
summary_DEG_3
```

###############################################################################


Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Past_Host_DE_genes_OO_OC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-OC'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_4 <- summary_DEG
summary_DEG_4
```


###############################################################################

# Effect of overall low pH vs. ambient pH
*However, no comparison of sites*

Second contrast

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Past_Host_DE_genes_OO_LC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-LC'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_5 <- summary_DEG
summary_DEG_5
```


###############################################################################

# DEG summary

```{r}
summary2 <- rbind(summary_DEG_1, summary_DEG_2, summary_DEG_3, summary_DEG_4)

summary2
```


```{r}
summary2$Destination <- c("Ambient", "Ambient", "Low pH", "Low pH", "Ambient vs. Low pH", "Ambient vs. Low pH", "Ambient vs. Low pH", "Ambient vs. Low pH") # "Ojo vs. Control", "Ojo"

summary2$Destination <- as.factor(summary2$Destination)
summary2$Destination <- factor(summary2$Destination, levels = c("Ambient", "Low pH", "Ambient vs. Low pH"))

# levels(summary2$group)
# "OC-LC" "OO-LO" "LO-LC" "OO-OC"

summary2$group <- revalue(summary2$group, c("OC-LC" = "Lagoon vs. Ojo", "OO-LO" = "Lagoon vs. Ojo", "LO-LC" = "Lagoon", "OO-OC" = "Ojo")) 

summary2
```


```{r}
#levels(summary2$group)
# "OC-LC" "OO-LO" "LO-LC" "OO-OC"
summary2$group <- factor(summary2$group, levels = c("OC-LC", "OO-LO", "LO-LC", "OO-OC"))

#write.csv(file = "DEG_summary_Past_Host.csv", x = summary2)
```


# Combined barplot
```{r}
plot.combo <- ggplot(summary2, aes(x = group, fill = LFC)) + 
  facet_wrap(~Destination, scales = "free_x") +
  geom_bar(data=subset(summary2, LFC=="Up"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_bar(data=subset(summary2, LFC=="Down"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_text(data=subset(summary2, LFC=="Up"), aes(group, n, label = n), position = position_dodge(width=0.9), vjust = -0.5) +
  geom_text(data=subset(summary2, LFC=="Down"), aes(group, n, label = -n), position = position_dodge(width=0.9), vjust = 1.5) +
  scale_fill_manual("LFC", values=c("black", "#56B4E9"), labels=c("Down-regulated", "Up-regulated"), guide=guide_legend(reverse=TRUE)) + #colorblind friendly
  ylab("Differentially expressed genes") +
  xlab("Native") + 
  scale_y_continuous(limits = c(-260, 160), breaks = seq(-250, 150, by = 50)) +
  theme_classic() +
  theme(legend.position = c(0.16, 0.18), legend.background = element_blank(), panel.background = element_rect(colour = "black"), axis.text.x = element_text(vjust = 0.5)) 

# axis.text.x   angle = 45,

# bottom right embedded inside plot c(0.84, 0.2)

#legend.position = "bottom",
#axis.text.x=element_text(vjust=0.5, size=10, face="italic"), 

plot.combo
```

```{r}
# save plot
ggsave("Past_Host_DEG_barplot_Low-pH-label.pdf", plot = plot.combo,  width = 5, height = 3, units = "in", dpi = 1200) #width = 6.5, height = 4.5
```


# Barplot without facet wrap
```{r}
plot.combo <- ggplot(summary2, aes(x = group, fill = LFC)) + 
  geom_bar(data=subset(summary2, LFC=="Up"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_bar(data=subset(summary2, LFC=="Down"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_text(data=subset(summary2, LFC=="Up"), aes(group, n, label = n), position = position_dodge(width=0.9), vjust = -0.5) +
  geom_text(data=subset(summary2, LFC=="Down"), aes(group, n, label = -n), position = position_dodge(width=0.9), vjust = 1.5) +
  scale_fill_manual("LFC", values=c("black", "#56B4E9"), labels=c("Down-regulated", "Up-regulated"), guide=guide_legend(reverse=TRUE)) + #colorblind friendly
  ylab("Differentially expressed genes") +
  xlab("Transplantation") + 
  scale_y_continuous(limits = c(-650, 620), breaks = seq(-600, 600, by = 200)) +
  theme_classic() +
  theme(legend.position = c(0.2, 0.18),  panel.background = element_rect(colour = "black"), axis.text.x = element_text(angle = 45, vjust = 0.5)) 

#legend.position = "bottom",
#axis.text.x=element_text(vjust=0.5, size=10, face="italic"), 

plot.combo
```


```{r}
# save plot
ggsave("Past_Host_DEG_barplot_no-facetwrap.pdf", plot = plot.combo,  width = 5, height = 4, units = "in", dpi = 1200) #width = 6.5, height = 4.5
```


## Session Info
```{r}
sessionInfo()
```


