---
title: "Venn_Past_Host"
author: "Veronica Radice"
date: "19/04/2021"
output: html_document
---

# Venn diagrams
## Differentially expressed genes

## Coral Host
## *Porites astreoides*


### Load packages
```{r setup, include=FALSE}
library(VennDiagram)
```

### Set working directory
```{r}
setwd("~/Documents/Rprojects/postdoc\ Rprojects/ODU_postdoc_Rprojects/Paper_Ojo_gene-expression/Porites_astreoides/host/")
```


# Import data

Summary
```{r}
deg <- read.csv("DEG_summary_Past_Host.csv", header = TRUE)
deg <- deg[,2:4]
deg
```

## Effect of Origin - corals transplanted to same Destination

Absolute number of DEG adjusted p-value < 0.1

## Lagoon-Ojo vs. Lagoon-Control

```{r}
de.genes_LO_LC <- readRDS("Past_Host_DE_genes_LO_LC_padj.1.rds")
de.genes_LO_LC <- as.data.frame(de.genes_LO_LC)
de.genes_LO_LC$group <- 'LO_LC'
de.genes_LO_LC$group <- as.factor(de.genes_LO_LC$group)

de.genes_LO_LC$expression <- ifelse(de.genes_LO_LC$log2FoldChange>0, 'up', 'down')
de.genes_LO_LC$expression <- as.factor(de.genes_LO_LC$expression)
dim(de.genes_LO_LC)
```

```{r}
LO_LC.up <- subset(de.genes_LO_LC, expression == "up")
dim(LO_LC.up)

LO_LC.down <- subset(de.genes_LO_LC, expression == "down")
dim(LO_LC.down)
```


## Ojo-Ojo vs. Lagoon-Ojo

```{r}
de.genes_OO_LO <- readRDS("Past_Host_DE_genes_OO_LO_padj.1.rds")
de.genes_OO_LO <- as.data.frame(de.genes_OO_LO)
de.genes_OO_LO$group <- 'OO_LO'
de.genes_OO_LO$group <- as.factor(de.genes_OO_LO$group)

de.genes_OO_LO$expression <- ifelse(de.genes_OO_LO$log2FoldChange>0, 'up', 'down')
de.genes_OO_LO$expression <- as.factor(de.genes_OO_LO$expression)
dim(de.genes_OO_LO)
```

```{r}
OO_LO.up <- subset(de.genes_OO_LO, expression == "up")
dim(OO_LO.up)

OO_LO.down <- subset(de.genes_OO_LO, expression == "down")
dim(OO_LO.down)
```


## Ojo-Ojo vs. Ojo-Control

```{r}
de.genes_OO_OC <- readRDS("Past_Host_DE_genes_OO_OC_padj.1.rds")
de.genes_OO_OC <- as.data.frame(de.genes_OO_OC)
de.genes_OO_OC$group <- 'OO_OC'
de.genes_OO_OC$group <- as.factor(de.genes_OO_OC$group)

de.genes_OO_OC$expression <- ifelse(de.genes_OO_OC$log2FoldChange>0, 'up', 'down')
de.genes_OO_OC$expression <- as.factor(de.genes_OO_OC$expression)
dim(de.genes_OO_OC)
```

```{r}
OO_OC.up <- subset(de.genes_OO_OC, expression == "up")
dim(OO_OC.up)

OO_OC.down <- subset(de.genes_OO_OC, expression == "down")
dim(OO_OC.down)
```


## Ojo-Control vs. Lagoon-Control

```{r}
de.genes_OC_LC <- readRDS("Past_Host_DE_genes_OC_LC_padj.1.rds")
de.genes_OC_LC <- as.data.frame(de.genes_OC_LC)
de.genes_OC_LC$group <- 'OC_LC'
de.genes_OC_LC$group <- as.factor(de.genes_OC_LC$group)

de.genes_OC_LC$expression <- ifelse(de.genes_OC_LC$log2FoldChange>0, 'up', 'down')
de.genes_OC_LC$expression <- as.factor(de.genes_OC_LC$expression)
dim(de.genes_OC_LC)
```

```{r}
OC_LC.up <- subset(de.genes_OC_LC, expression == "up")
dim(OC_LC.up)

OC_LC.down <- subset(de.genes_OC_LC, expression == "down")
dim(OC_LC.down)
```


##############################################################################

# All DEG - 4-way venn comparing Control and Ojo destination pairwise contrasts
```{r}
venn.diagram(x = list(de.genes_LO_LC$gene, de.genes_OO_LO$gene, de.genes_OO_OC$gene, de.genes_OC_LC$gene), 
             category.names = c("Lagoon-Ojo vs Lagoon-Control", "Ojo-Ojo vs Lagoon-Ojo", "Ojo-Ojo vs Ojo-Control", "Ojo-Control vs Lagoon-Control"),
         filename = 'Past_Host_venn_4-way.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 6000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3"), #, "goldenrod"
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        #cat.pos = c(-27, 27, 135, 135, 27),
        #cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "Porites astreoides",
        scaled = FALSE)

#cat.pos = c(0, 0),
```

## Upregulated
```{r}
venn.diagram(x = list(LO_LC.up$gene, OO_LO.up$gene, OO_OC.up$gene, OC_LC.up$gene), 
             category.names = c("Lagoon-Ojo vs Lagoon-Control", "Ojo-Ojo vs Lagoon-Ojo", "Ojo-Ojo vs Ojo-Control", "Ojo-Control vs Lagoon-Control"),
         filename = 'Past_Host_venn_up_labels.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 6000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3"), #, "goldenrod"
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        #cat.pos = c(-27, 27, 135, 135, 27),
        #cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "",
        scaled = FALSE)

#cat.pos = c(0, 0),
```


## Upregulated - without labels
```{r}
venn.diagram(x = list(LO_LC.up$gene, OO_LO.up$gene, OO_OC.up$gene, OC_LC.up$gene), 
             category.names = c("", "", "", ""),
         filename = 'Past_Host_venn_up.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 6000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3"), #, "goldenrod"
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        #cat.pos = c(-27, 27, 135, 135, 27),
        #cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "",
        scaled = FALSE)

#cat.pos = c(0, 0),
```


## Downregulated - 3-way because one contrast has zero downregulated DEG
### With labels
```{r}
venn.diagram(x = list(LO_LC.down$gene, OO_LO.down$gene, OO_OC.down$gene), 
             category.names = c("Lagoon-Ojo vs Lagoon-Control", "Ojo-Ojo vs Lagoon-Ojo", "Ojo-Ojo vs Ojo-Control"),
         filename = 'Past_Host_venn_down_labels.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 6000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1"), #, "goldenrod"
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        #cat.pos = c(-27, 27, 135, 135, 27),
        #cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "",
        scaled = FALSE)

#cat.pos = c(0, 0),
```


## Downregulated - 3-way because one contrast has zero downregulated DEG
```{r}
venn.diagram(x = list(LO_LC.down$gene, OO_LO.down$gene, OO_OC.down$gene), 
             category.names = c("", "", ""),
         filename = 'Past_Host_venn_down.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 6000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1"), #, "goldenrod"
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        #cat.pos = c(-27, 27, 135, 135, 27),
        #cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "",
        scaled = FALSE)

#cat.pos = c(0, 0),
```


# All DEG - 3-way venn comparing Control and Ojo destination pairwise contrasts
```{r}
venn.diagram(x = list(de.genes_LO_LC$gene, de.genes_OO_LO$gene, de.genes_OO_OC$gene), 
             category.names = c("Lagoon-Ojo vs Lagoon-Control", "Ojo-Ojo vs Lagoon-Ojo", "Ojo-Ojo vs Ojo-Control"),
         filename = 'Past_Host_venn_3-way.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4500, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1"), #"deeppink3" , "goldenrod"
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 0),
        cat.dist = c(0.055, 0.055, 0.065),
        main = "Porites astreoides",
        scaled = FALSE)

#cat.pos = c(0, 0),
```




### Session Info
```{r}
sessionInfo()
```

